---
title: '<font size = 7 color = "white">Comparing Potential Sampling Plans.</font>'
subtitle: 
author: '<font size = 5 color = "white"> Kyle Weber and Ian Vanwright </font>'
institute: '<font size = 6 color = "white">West Chester University of Pennsylvania</font><br> '
date: ' <font color = "white" size =4> Prepared for<br> </font> <font color = "gold" size = 6> STA490: Data Visualization </font> <br><br> <font color = "white" size = 3> Slides available at: https://rpubs.com/KW986324 AND https://github.com/Kyle-Weber/STA490</font>'
output:
  xaringan::moon_reader:
    css: xaringan-themer01.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("pander")) {
   install.packages("pander")
   library(pander)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
knitr::opts_chunk$set(
                  fig.width=3, 
                  fig.height=3, 
                  fig.retina=12,
                  out.width = "100%",
                  cache = FALSE,
                  echo = FALSE,
                  message = FALSE, 
                  warning = FALSE,
                  hiline = TRUE
                  )
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
  style_duo_accent(primary_color = "#1F4257",
          secondary_color = "#380F2A",
          # fonts
          header_font_google = google_font("Martel"),
          text_font_google = google_font("Lato"),
          code_font_google = google_font("Fira Mono"))
  

```


```{r, include=FALSE}
 if (!require("xaringanthemer")) {
      install.packages("xaringanthemer")
      library(xaringanthemer)
     }
```




```{r, include=FALSE}
## Data Set
library(dplyr)
library(ggplot2)

github_urls <- c(
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational01.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational02.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational03.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational04.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational05.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational06.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational07.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational08.csv",
  "https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational09.csv"
)

BankLoan <- read.csv("https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational01.csv", header = TRUE)


for (i in 2:9) {
  file_url <- paste0("https://raw.githubusercontent.com/Kyle-Weber/STA490/main/data%20sets/w06-SBAnational0", i, ".csv")
  data_to_append <- read.csv(file_url, header = TRUE)
  BankLoan <- rbind(BankLoan, data_to_append)
}


```


class: top, center
##    Table of Contents: 

.left[

- Data Cleaning

- Combining Categories

- Default Rate

- Analyzing Four Sampling Plans

- Comparing Default Rates

- Default Rate Visualization

- Conclusion
]






---


class: top, center
##    Data Cleaning: 


.left[


- Removed All Missing Values 

- Changed Currency Values to Numerical

- Remaining Observations is 897154


]


```{r, include=FALSE}
## Step 1: Deleting NA Values


library(dplyr)

# check values
unique_values <- lapply(BankLoan[, c("DisbursementGross", "BalanceGross", "ChgOffPrinGr", "GrAppv", "SBA_Appv")], unique)


miss1 <- BankLoan[which(BankLoan$MIS_Status == ""), ]
nrow(miss1)

miss2 <- BankLoan[which(BankLoan$State == ""), ]
nrow(miss2)


BankLoan <- BankLoan[!(is.na(BankLoan$State) | BankLoan$State == ""), ]


missTotal <- rbind(miss1, miss2)
nrow(missTotal)

BankLoan_cleaned <- BankLoan[-which(BankLoan$State == ""), ]
BankLoan_cleaned <- BankLoan[-which(BankLoan$MIS_Status == ""), ]



```



---

class: top, center
##    Combining Categories: 

.left[

- Variable "State" Combined into 5 Regions

- Represent Geographical Region

- Frequency Table of Regions

- Lowest is SouthWest

]




```{r, include=FALSE}
##  Step 2: Choosing Categorical Variable

BankLoan_cleaned <- BankLoan_cleaned %>%
  mutate(State = case_when(
    State %in% c("AL", "AR", "GA", "KY", "LA", "MS", "SC", "TN", "WV", "OK") ~ "SW",
    State %in% c("AK", "AZ", "CA", "CO", "HI", "ID", "MT", "NV", "NM", "OR", "UT", "WA", "WY") ~ "W",
    State %in% c("CT", "DE", "MA", "MD", "ME", "NH", "NJ", "NY", "PA", "RI", "VT") ~ "NE",
    State %in% c("IL", "IN", "IA", "KS", "MI", "MN", "MO", "NE", "ND", "OH", "SD", "WI") ~ "MW",
    State %in% c("FL", "NC", "TX", "VA", "DC") ~ "SE",
    TRUE ~ as.character(State)
  ))
```

```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)

State_frequencies <- BankLoan_cleaned %>%
  count(State)

# Print the result as a table
kable(State_frequencies, col.names = c("Region", "Frequency"))



```


---
class: top, center

##    Calculating Default Rates:

.left[

- Original Rates for Data Set

- Provides a Comparison to the Sampling Plans

-  Southeast has the highest rate
]

<!--If need time can talk about why they are important-->

```{r, echo=FALSE, warning=FALSE}
##  Step 3: Calculating Default Rates

default_rates <- BankLoan_cleaned %>%
  group_by(State) %>%
  summarise(Default_Rate = mean(MIS_Status == "CHGOFF"))


kable(default_rates, col.names= c("State", "Default Rate"))

```


---

```{r, echo=FALSE, warning=FALSE}
clusters <- unique(BankLoan_cleaned$Zip)

```

class: top, center

##    Simple Random Sampling Process:

.left[

- Random Sample of 4000

<!-- why 4000 less than 5% of original data set -->

- Taken from Cleaned Data Set

- Frequency  and Default Rate Table

]


```{r, echo=FALSE, warning=FALSE}
random_sample <- BankLoan_cleaned[sample(nrow(BankLoan_cleaned), 4000), ]


random_sample_dim <- dim(random_sample)

size_var_count <- data.frame(Size = random_sample_dim[1], Var.count = random_sample_dim[2])

kable(t(size_var_count), col.names = c("Size", "Var.count"))








# Define a function to calculate default rates for each region
calculate_region_default_rates <- function(data) {
  # Create a data frame to store the results
  region_default_rates <- data.frame(Region = character(), Sample_Default_Rate = numeric(), stringsAsFactors = FALSE)
  
  # Calculate default rates for each region
  for (region in unique(data$State)) {
    region_data <- data[data$State == region, ]
    region_default_rate <- sum(region_data$MIS_Status == "CHGOFF") / nrow(region_data)
    region_default_rates <- rbind(region_default_rates, data.frame(Region = region, Sample_Default_Rate = region_default_rate))
  }
  
  return(region_default_rates)
}

# Calculate default rates for each region within the state for Simple Random Sampling
simple_random_region_default_rates <- calculate_region_default_rates(random_sample)

# Print the default rates for each region for Simple Random Sampling
kable(simple_random_region_default_rates)


```

---



class: top, center
##    Systematic Sample:

.left[

- Jump size created on 4000 observations

- Selects observations from fixed intervals

- Representative of entire population

- Rounding Error

]


```{r, echo=FALSE, warning=FALSE}
jump.size = dim(BankLoan_cleaned)[1] %/% 4000
rand.starting.pt = sample(1:jump.size, 1)
sampling.id = seq(rand.starting.pt, dim(BankLoan_cleaned)[1], jump.size)
sys.sample = BankLoan_cleaned[sampling.id, ]

sys.Sample.dim = dim(sys.sample)
names(sys.Sample.dim) = c("Size", "Var.count")
kable(t(sys.Sample.dim))







# Calculate default rates for each region within the state for Systematic Sampling
systematic_region_default_rates <- calculate_region_default_rates(sys.sample)

# Print the default rates for each region for Systematic Sampling
kable(systematic_region_default_rates)


```


---

class: top, center
##    Stratified Sampling: 

.left[


- Strata Size for Regions

- 4000 Observations

- Default Rates


]

```{r, echo=FALSE, warning=FALSE}
library(kableExtra)
freq.table = table(BankLoan_cleaned$State)  # frequency table of strNAICS
rel.freq = freq.table/sum(freq.table)   # relative frequency 
strata.size = round(rel.freq*4000)      # strata size allocation
strata.names=names(strata.size)  

kable(t(strata.size))

```



```{r, echo=FALSE, warning=FALSE}
strata.sample = BankLoan_cleaned[1,]    # create a reference data frame
strata.sample$add.id = 1   # add a temporary ID to because in the loop
                           # i =2 testing a single iteration
for (i in 1:length(strata.names)){
   ith.strata.names = strata.names[i]   # extract data frame names
   ith.strata.size = strata.size[i]     # allocated stratum size
   # The following code identifies observations to be selected
   ith.sampling.id = which(BankLoan_cleaned$State==ith.strata.names) 
   ith.strata = BankLoan_cleaned[ith.sampling.id,]  # i-th stratified population
   ith.strata$add.id = 1:dim(ith.strata)[1]  # add sampling list/frame
   # The following code generates a subset of random ID
   ith.sampling.id = sample(1:dim(ith.strata)[1], ith.strata.size) 
   ## Create a selection status -- pay attention to the operator: %in% 
   ith.sample =ith.strata[ith.strata$add.id %in%ith.sampling.id,]
   ## dim(ith.sample)         $ check the sample
   strata.sample = rbind(strata.sample, ith.sample)  # stack all data frame!
 }
 # dim(strata.sample)
 strat.sample.final = strata.sample[-1,]  # drop the 

 
 
 

 
 
# Calculate default rates for each region within the state for Stratified Sampling
stratified_region_default_rates <- calculate_region_default_rates(strat.sample.final)

# Print the default rates for each region for Stratified Sampling
kable(stratified_region_default_rates)


```


---

class: top, center
## Cluster Sampling:

.left[

- ZipCode used to Define Clusters

- 140 Clusters Taken



]


```{r, echo=FALSE, warning=FALSE}
clusters <- unique(BankLoan_cleaned$Zip)

```



```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(kableExtra)

clusters <- unique(BankLoan_cleaned$Zip)

# Take a cluster sample using zip codes
selected_clusters <- sample(clusters, 140)  # Adjust the number of clusters
cluster_sample <- BankLoan_cleaned[BankLoan_cleaned$Zip %in% selected_clusters, ]


# Print the size and variable count
cluster_sample_dim <- dim(cluster_sample)
size_var_count_cluster <- data.frame(Size = cluster_sample_dim[1], Var.count = cluster_sample_dim[2])
kable(t(size_var_count_cluster), col.names = c("Size", "Var.count"))



# Calculate default rates 
clustering_region_default_rates <- cluster_sample %>%
  group_by(State) %>%
  summarise(Default_Rate = mean(MIS_Status == "CHGOFF"))

# Print default rates for each region
kable(clustering_region_default_rates, col.names = c("Region", "Default Rate"))

```





---

class: top, center
##   Default Rate Comparison:

.left[
- Comparison of Default Rates Across Regions

- Sample Default Rate vs True Default Rate

- Systematic or Stratified
]

```{r interactive-logistic-regression-summary, echo=FALSE, fig.cap="Summary of inferential statistics of the full model", out.width="100%"}
# Define a function to calculate default rates for each region
calculate_region_default_rates <- function(data) {
  # Create a data frame to store the results
  region_default_rates <- data.frame(Region = character(), Sample_Default_Rate = numeric(), Subpopulation_Default_Rate = numeric(), stringsAsFactors = FALSE)
  
  # Calculate default rates for each region
  for (region in unique(data$State)) {
    region_data <- data[data$State == region, ]
    region_sample_default_rate <- sum(region_data$MIS_Status == "CHGOFF") / nrow(region_data)
    region_subpopulation_default_rate <- sum(BankLoan_cleaned$MIS_Status[BankLoan_cleaned$State == region] == "CHGOFF") / sum(BankLoan_cleaned$State == region)
    region_default_rates <- rbind(region_default_rates, data.frame(Region = region, Sample_Default_Rate = region_sample_default_rate, Subpopulation_Default_Rate = region_subpopulation_default_rate))
  }
  
  return(region_default_rates)
}

# Calculate default rates for each region within the state for each sampling process
simple_random_region_default_rates <- calculate_region_default_rates(random_sample)
systematic_region_default_rates <- calculate_region_default_rates(sys.sample)
clustering_region_default_rates <- calculate_region_default_rates(cluster_sample)
stratified_region_default_rates <- calculate_region_default_rates(strat.sample.final)

# Combine default rates for each region from all sampling processes into a single data frame
all_region_default_rates <- rbind(
  data.frame(Sampling_Plan = "Simple Random", simple_random_region_default_rates),
  data.frame(Sampling_Plan = "Systematic", systematic_region_default_rates),
  data.frame(Sampling_Plan = "Clustering", clustering_region_default_rates),
  data.frame(Sampling_Plan = "Stratified", stratified_region_default_rates)
)



library(DT)

# Create a summary table
summary_table <- all_region_default_rates

# Round numeric columns to 5 decimal places
summary_table$Sample_Default_Rate <- round(summary_table$Sample_Default_Rate, digits = 5)
summary_table$Subpopulation_Default_Rate <- round(summary_table$Subpopulation_Default_Rate, digits = 5)

# Render the summary table interactively
datatable(summary_table, 
          options = list(dom = 't', 
                         pageLength = 61, 
                         scrollY = '200px', 
                         scrollCollapse = TRUE))



```


---

class: top, center

##    Comparison Between Each other:

.left[
- Default Rate for each Region Based on Sampling Plan
]

```{r all_region_default_rates,  echo=FALSE, fig.asp=0.5, fig.align='center'}
# Load required libraries
# Load required libraries
library(ggplot2)

# Define column names for region default rates
col_names <- c("Region", "Default_Rate")

# Set column names for each data frame
names(simple_random_region_default_rates) <- col_names
names(systematic_region_default_rates) <- col_names
names(clustering_region_default_rates) <- col_names
names(stratified_region_default_rates) <- col_names

# Add the Sampling_Plan column to each data frame
simple_random_region_default_rates$Sampling_Plan <- "Simple Random"
systematic_region_default_rates$Sampling_Plan <- "Systematic"
clustering_region_default_rates$Sampling_Plan <- "Clustering"
stratified_region_default_rates$Sampling_Plan <- "Stratified"

# Combine all region default rates into a single data frame
all_region_default_rates <- rbind(simple_random_region_default_rates, 
                                  systematic_region_default_rates, 
                                  clustering_region_default_rates, 
                                  stratified_region_default_rates)

# Create a bar plot with adjusted font sizes and custom colors
ggplot(all_region_default_rates, aes(x = Region, y = Default_Rate, fill = Sampling_Plan)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Simple Random" = "orange", 
                                "Systematic" = "darkgreen", 
                                "Clustering" = "blue", 
                                "Stratified" = "magenta")) +  # Set custom colors
  labs(title = "Default Rates for Each Region by Sampling Plan",
       x = "Region", y = "Default Rate", fill = "Sampling Plan") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6),      # Adjust x-axis tick label font size
        axis.text.y = element_text(size = 4),      # Adjust y-axis tick label font size
        axis.title.x = element_text(size = 4),    # Adjust x-axis label font size
        axis.title.y = element_text(size = 4),    # Adjust y-axis label font size
        legend.text = element_text(size = 6),      # Adjust legend text font size
        legend.title = element_text(size = 6),    # Adjust legend title font size
        plot.title = element_text(size = 6))      # Adjust plot title font size


```



---

class: top, center

##    Visualization For Final Choice:


```{r plot_1,  echo=FALSE, fig.asp=0.5, fig.align='center'}

# Separate data frames for each pair of sampling plans
summary_table_1 <- subset(summary_table, Sampling_Plan %in% c("Simple Random", "Systematic"))
summary_table_2 <- subset(summary_table, Sampling_Plan %in% c("Clustering", "Stratified"))

# Plot for the first pair of sampling plans
plot_1 <- ggplot(summary_table_1, aes(x = Region, y = Sample_Default_Rate, fill = Sampling_Plan)) +
  geom_col(position = "dodge", color = "black", width = 0.5) +
  geom_col(aes(y = Subpopulation_Default_Rate), position = "dodge", alpha = 0.5, color = "black", width = 0.5) +
  labs(title = "Default Rates by Region",
       y = "Default Rate",
       fill = "Sampling Plan") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 4),  # Rotate and decrease x-axis labels
        axis.text.y = element_text(size = 3),     # Decrease y-axis tick label font size
        axis.title.y = element_text(size = 3),    # Decrease y-axis label font size
        legend.text = element_text(size = 3),     # Decrease legend text font size
        legend.title = element_text(size = 3),    # Decrease legend title font size
        plot.title = element_text(size = 3),      # Decrease plot title font size
        strip.text.x = element_text(size = 4)) +  # Decrease font size of top x-axis labels
  facet_wrap(~ Sampling_Plan, scales = "free_x")

plot_1



```


---

class: top, center

##    Visualization for Final Choice






```{r plot_2,  echo=FALSE, fig.asp=0.5, fig.align='center'}
# Plot for the second pair of sampling plans
plot_2 <- ggplot(summary_table_2, aes(x = Region, y = Sample_Default_Rate, fill = Sampling_Plan)) +
  geom_col(position = "dodge", color = "black", width = 0.5) +
  geom_col(aes(y = Subpopulation_Default_Rate), position = "dodge", alpha = 0.5, color = "black", width = 0.5) +
  labs(title = "Comparison of Sample Default Rates to Subpopulation Default Rates by Region",
       y = "Default Rate",
       fill = "Sampling Plan") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 4),  # Rotate and decrease x-axis labels
        axis.text.y = element_text(size = 3),     # Decrease y-axis tick label font size
        axis.title.y = element_text(size = 3),    # Decrease y-axis label font size
        legend.text = element_text(size = 3),     # Decrease legend text font size
        legend.title = element_text(size = 3),    # Decrease legend title font size
        plot.title = element_text(size = 4),      # Decrease plot title font size
        strip.text.x = element_text(size = 4)) +  # Decrease font size of top x-axis labels
  facet_wrap(~ Sampling_Plan, scales = "free_x")



plot_2


```












---
class: top, center

##     Conclusion

.left[
- Final Choice

- Why this is important

- Questions
]















